Chemical tagging: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-15 papers/clusters.tex
figures
figure tweaks
k-means

% figure shit
\newcommand{\insanefigure}[1]{%
\includegraphics[width=0.33\textwidth]{./figs/#1_MgFe.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_NaO.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_CN.png}\\
\includegraphics[width=0.33\textwidth]{./figs/#1_AlMg.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_SAl.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_MgK.png}\\
\includegraphics[width=0.33\textwidth]{./figs/#1_sky.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_decv.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_HR.png}}

@@

\title{Chemical tagging: \\
       Identification of stellar phase-space structures \\
       purely by chemical-abundance similarity}
\author{David~W.~Hogg\altaffilmark{1,2,3,4},
        Andrew~R.~Casey\altaffilmark{5},
        M.~Ness\altaffilmark{4},
        Hans-Walter~Rix\altaffilmark{4}, \&
        Daniel~Foreman-Mackey\altaffilmark{6}}
\altaffiltext{1}{SCDA}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Phyics,
  New York University, 4 Washington Pl., room 424, New York, NY, 10003, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th Floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{5}{IoA}
\altaffiltext{6}{UW}
\email{david.hogg@nyu.edu}

\begin{abstract}
% Context:
The chemical-tagging hypothesis is that every star ought to carry---%
in its detailed chemical abundances---%
an informative signature of its molecular-cloud birth location and time.
This idea has not previously yielded scientific successes, probably because of the
noise in chemical-abundance measurements.
However, we have succeeded in improving vastly spectroscopic measurements with \thecannon\ 
(now delivering individual abundances with precisions around 0.04-dex).
% Aims:
We test the chemical-tagging hypothesis by looking at clusters in abundance space
and confirming that they are clustered in phase space.
% Methods:
We identify (by the k-means algorithm) overdensities of stars observed by the \apogee\ spectroscopic survey
in a 15-dimensional chemical-abundance space delivered by \thecannon,
and plot the associated stars in phase space.
We use \emph{only} abundance-space information (no positional information) to identify stellar groups.
% Results:
We find that clusters in abundance space are indeed clusters in phase space.
We recover some known clusters and find other interesting structures.
This confirms the chemical-tagging hypothesis and verifies the precision of abundances delivered by \thecannon.
This is the first-ever project to identify phase-space structures by blind search in abundance space;
the prospects for future data sets are very good.
\end{abstract}

@@

Although only hints of small-scale clustering in the abundance space
are visible in \figurename~\ref{fig:all}, exploration of the data indicates that
known clusters do appear in the abundance space as over-densities.
This encourages us to look for over-densities automatically in the
abundance space and see if anything found that way would be over-dense
in phase space.
The simplest method for clustering points in $D$-dimensional space is
the \emph{k-means} algorithm (DWH CITE).
This algorithm is fast and performs well in practice in problems of
this nature.

DWH: Scaling of the data
DWH: number of clusters
DWH: restarts for local minima
DWH: Citation out to numpy and sklearn.

@@

@@

DWH: For the first time, we see small-scale informative features in abundance space.

DWH: When we identify the small-scale features in abundance space, they are informative in position space.

>>>>

% figure shit
\newcommand{\insanefigure}[1]{%
\includegraphics[width=0.33\textwidth]{./figs/#1_MgFe.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_MgK.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_AlMg.png}\\
\includegraphics[width=0.33\textwidth]{./figs/#1_NaO.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_CN.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_SAl.png}\\
\includegraphics[width=0.33\textwidth]{./figs/#1_sky.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_decv.png}%
\includegraphics[width=0.33\textwidth]{./figs/#1_HR.png}}

@@

\title{Chemical tagging: \\
       Identification of stellar phase-space structures \\
       purely by chemical-abundance similarity}
\author{David~W.~Hogg\altaffilmark{1,2,3,4},
        Andrew~R.~Casey\altaffilmark{5},
        M.~Ness\altaffilmark{4},
        Hans-Walter~Rix\altaffilmark{4},
    and~Daniel~Foreman-Mackey\altaffilmark{6}}
\altaffiltext{1}{SCDA}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Phyics,
  New York University, 4 Washington Pl., room 424, New York, NY, 10003, USA}
\altaffiltext{3}{Center for Data Science, New York University, 726 Broadway, 7th Floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{5}{IoA}
\altaffiltext{6}{UW}
\email{david.hogg@nyu.edu}

\begin{abstract}
% Context:
The chemical-tagging hypothesis is that every star ought to carry---%
in its detailed chemical abundances---%
an informative signature of its molecular-cloud birth location and time.
This idea has not previously yielded scientific successes, probably because of the
noise in chemical-abundance measurements.
However, we have succeeded in improving vastly spectroscopic measurements with \thecannon\ 
(now delivering individual abundances with precisions around 0.04-dex).
% Aims:
We test the chemical-tagging hypothesis by looking at clusters in abundance space
and confirming that they are clustered in phase space.
% Methods:
We identify (by the k-means algorithm) overdensities of stars observed by the \apogee\ spectroscopic survey
in a 15-dimensional chemical-abundance space delivered by \thecannon,
and plot the associated stars in phase space.
We use \emph{only} abundance-space information (no positional information) to identify stellar groups.
% Results:
We find that clusters in abundance space are indeed clusters in phase space.
We recover some known clusters and find other interesting structures.
This confirms the chemical-tagging hypothesis and verifies the precision of abundances delivered by \thecannon.
This is the first-ever project to identify phase-space structures by blind search purely in abundance space;
the prospects for future data sets are very good.
\end{abstract}

@@

Although only hints of small-scale clustering in the abundance space
are visible in \figurename~\ref{fig:all}, exploration of the data indicates that
known clusters do appear in the abundance space as over-densities.
This encourages us to look for over-densities automatically in the
abundance space and see if anything found that way would be over-dense
in phase space.
The simplest method for clustering points in $D$-dimensional space is
the \emph{k-means} algorithm (DWH CITE).
Briefly, the k-means algorithm is to find the points closest (in
$D$-space) to each of $K$ centers, and then take the mean of those
points to update the locations of the centers; iterate to convergence.
The output is the locations of the $K$ centers and the assignments of
all points to centers.
This algorithm is fast and performs well in practice in problems of
this nature.

The k-means algorithm has disadvantages, of course.
One is that $K$ must be chosen by hand (or heuristically at best).
Here we are only demonstrating a concept; we don't need to have the
best possible clustering.
For this reason we simply choose $K=128, 256, 512$ and look at all the
results.
The algorithm also only performs local optimization.
At each $K$ we perform 32 restarts with different initializations, and
preserve the best clustering (best according to the k-means score).

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate measurement
precisions obtained by \thecannon\ before running k-means.
This makes the space close to isotropic in measurment uncertainty or
observational precision.

When the k-means results are returned, we construct the $D\times D$ empirical variance tensor for the DWH THIS IS WRONG!

DWH: Citation out to numpy and sklearn.

@@

\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Whatevsinator.\label{fig:infallcluster}}
\end{figure}

\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Whatevsinator.\label{fig:Pal5}}
\end{figure}

\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Whatevsinator.\label{fig:Sgr}}
\end{figure}

\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Whatevsinator.\label{fig:halo}}
\end{figure}

\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Whatevsinator.\label{fig:disk}}
\end{figure}

@@

DWH: For the first time/H], [we see small-scale informative features in abundance space.

DWH: When we identify the small-scale features in abundance space,they are informative in position space.
